### 2023-12-05 main.tex
ll

Observing now that $c_{x}(1-\delta_{\tau_{x},N})$ can be replaced by  $c_{x}-\delta_{\tau_{x},N}$, we finally have
\begin{equation}\label{elementsABS}
	\begin{split}
		\Psi^{'}(\bm{\tau})=\frac{1}{(L+1)!}\sum_{c_{1}=\delta_{\tau_{1},N}}^{1}\ldots\sum_{c_{L}=\delta_{\tau_{L},N}}^{1}\prod_{x=1}^{L}(\alpha_{\tau_{x}}-\beta_{\tau_{x}})^{1-c_{x}}\left(1+\sum_{j=x}^{L}c_{j}\right)\beta_{\tau_{x}}^{c_{x}-\delta_{\tau_{x},N}}\dt
	\end{split}
\end{equation} 
\begin{flushright}
    $\square$
\end{flushright}
\begin{remark} By changing the summation indices from $c_x\in \{\delta_{\tau_{x},N},1\}$ to $c_{x}^{'}\in \{0,1-\delta_{\tau_{x},N}\}$ and by  the fact that occupation of each site is bounded by $1$ we rewrite \eqref{ABS} as 
\begin{equation}\label{ABS_1}
	\Psi^{'}(\bm{\tau})=\sum_{c^{'}_{1}=0}^{1-\delta_{\tau_{1},N}}\ldots\sum_{c^{'}_{L}=0}^{1-\delta_{\tau_{L},N}}\frac{(L+1-\sum_{x=1}^{L}c^{'}_{x})!}{(L+1)!}\prod_{x=1}^{L}\left((\alpha_{\tau_{x}}-\beta_{\tau_{x}})\left(2+L-x-\sum_{j=x}^{L}c^{'}_{j}\right)\right)^{c^{'}_{x}}\beta_{\tau_{x}}^{(1-c^{'}_{x}-\delta_{\tau_{x},N})}
\end{equation}
where we used the fact that 
\begin{equation}
	\left(1+\sum_{j=x}^{L}c_{j}\right) =\left(2+L-x-\sum_{j=x}^{L}c^{'}_{j}\right)\dt
\end{equation}
Equation \eqref{ABS_1} will be useful in the following. 
\end{remark}
\subsubsection{Closed formula for correlations}\label{correlation-section}
We consider $m\in \{1,\ldots,L\}$ sites with coordinates $x_{1},\ldots,x_{m}\in \{1,\ldots,L\}$ such that $x_{k}<x_{k+1}$ for all $k\in \{1,\ldots,m-1\}$. We fix $a_{k}\in\{1,\ldots,N-1\}$ and we would like to compute
\begin{equation}
	\mathbb{E}\left[\prod_{k=1}^{m}\mathbbm{1}_{\{Y_{x_{k}}=a_{k}\}}\right]=\sum_{\bm{\tau}\in \Omega^{'}}\left(\prod_{k=1}^{m}\mathbbm{1}_{\{\tau_{x_{k}}=a_{k}\}}\right)\langle\bm{\tau}|\Psi\rangle\dt%=\sum_{\bm{\tau}\in \Omega^{'}}\left(\prod_{i=1}^{m}\mathbbm{1}_{\{\tau_{y_{i}}=s_{i}\}}\right)\Psi(\bm{\tau})
\end{equation}
We fix the dual configuration $\bm{\xi}=\sum_{k=1}^{m}\delta_{a_{k}}^{x_{k}}$, corresponding to a vector $|\bm{\xi}\rangle$, and we have 
\begin{equation}
	\prod_{k=1}^{m}\mathbbm{1}_{\{\tau_{x_{k}}=a_{k}\}}=D(\bm{\tau},\bm{\xi})\dt
\end{equation}
Thus it follows that 
\begin{align}
	\mathbb{E}\left[\prod_{k=1}^{m}\mathbbm{1}_{\{Y_{x_{k}}=a_{k}\}}\right]=&\sum_{\bm{\tau}\in \Omega^{'}}D(\bm{\tau},\bm{\xi})\langle\bm{\tau}|\Psi\rangle
	=\sum_{\bm{\tau}\in \Omega^{'}}\langle \bm{\tau}|D|\bm{\xi}\rangle\langle\bm{\tau}|\Psi\rangle\dt
\end{align}
Denoting by $|\hat{\bm{\xi}}\rangle=|\xi^{1},\ldots,\xi^{L}\rangle$, i.e. the vector constructed from $|\bm{\xi}\rangle$ by removing the components at sites $\{0\}$ and $\{L+1\}$ and recalling that  
\begin{equation}
	S_{1}^{T}=\prod_{x=1}^{L}d_{x}
\end{equation} 
we have
\begin{equation}
	\langle \bm{\tau}|S_{1}^{T}|\hat{\bm{\xi}}\rangle=\langle \bm{\tau}|D|\bm{\xi}\rangle \dt
\end{equation}
Therefore, it follows that, using the resolution of the identity
\begin{align}
\label{corr-psi1}
	\mathbb{E}\left[\prod_{k=1}^{m}\mathbbm{1}_{\{Y_{x_{k}}=a_{k}\}}\right]=& \sum_{\bm{\tau}\in \Omega^{'}}\langle \bm{\tau}|D|\bm{\xi}\rangle\langle\bm{\tau}|\Psi\rangle
	=\sum_{\bm{\tau}\in \Omega^{'}}\langle \bm{\tau}|S_{1}^{T}|\hat{\bm{\xi}}\rangle\langle\bm{\tau}|\Psi\rangle\nonumber\\
	=&\sum_{\bm{\tau}\in \Omega^{'}}\langle \hat{\bm{\xi}}|S_{1}|\bm{\tau}\rangle\langle\bm{\tau}|\Psi\rangle
	=\langle \hat{\bm{\xi}}|S_{1}|\Psi\rangle
	=\langle \hat{\bm{\xi}}|\Psi^{'}\rangle\dt
\end{align}

@@

{\color{blue}
\begin{remark}
Alternatively, by a direct computation we immediately verify 
\begin{equation}
\begin{split}
\langle N,\ldots,N,a_1,N,\ldots,N,a_m,N,\ldots,N|\Psi^{'}\rangle
 =\sum_{\bm{\tau}\in\Omega^{'}}\prod_{k=1}^{m}\delta_{{{\tau}_{x_k}},a_{k}}\langle \bm{\tau}|\Psi\rangle=\mathbb{E}\left[\prod_{k=1}^{m}\mathbbm{1}_{\{Y_{x_{k}}=a_{k}\}}\right]
 \end{split}
\end{equation} 
where the entries $a_i\neq N$ are at position $x_i$. Here we only  used  that $|\Psi^{'}\rangle=S_1|\Psi\rangle$ and
\begin{equation}
 \langle \tau |S_1=\begin{cases}
                    \sum_{\sigma=1}^N\langle \sigma|  \quad\text{if}\quad \tau=N\\
                    \langle \tau| \quad\text{else} 
                   \end{cases}\,.
\end{equation} 
\end{remark}

}
\todo{ok?}

\subsubsection{Interpretation of $\Psi^{'}(\bm{\tau})$ as marginals of the steady state distribution}\label{subsection-interpretationPSIp}
%We argue similarly as in the Section~\ref{correlation-section}. 
We still miss the interpretation of $\Psi^{'}(\bm{\tau})$ when  $\bm{\tau}= (N,\ldots,N)$. We show here that for this particular configuration
$\Psi^{'}(N,\ldots,N) =1$. Indeed, for any $\bm{\tau}\in \Omega^{'}$ we have that 
\begin{align}
	\Psi^{'}(\bm{\tau})=&\langle\bm{\tau}|\Psi^{'}\rangle=\sum_{\bm{\sigma}\in \Omega}\langle \bm{\tau}|S_{1}|\bm{\sigma}\rangle\langle \bm{\sigma}|\Psi\rangle=\sum_{\sigma\in \Omega}\langle \bm{\sigma}|S_{1}^{T}|\bm{\tau}\rangle\langle \bm{\sigma}|\Psi\rangle\dt
\end{align}
\todo{what was Omega again?}
We introduce the dual configuration $\bm{\xi}(\bm{\tau})=(N,\tau_{1},\ldots,\tau_{L},N)$ where two extra-sites occupied by the species $N$ have been attached to the vector $\bm{\tau}$, transforming it from a vector of dimension $L$ to one of dimension $L+2$. More explicitly, this dual configuration is given by
\begin{equation}
	\begin{cases}
	\xi_{a}^{x}(\bm{\tau})= \delta_{\tau_{x},a}\qquad \forall a\in \{1,\ldots N\}\quad \forall x\in \{1,\ldots,L\}\\
	\xi_{N}^{0}(\bm{\tau})=\xi_{N}^{L+1}(\bm{\tau})=1
\end{cases}\dt
\end{equation} Now we consider the elements $\langle\bm{\sigma}|D|\xi(\bm{\tau})\rangle$ of the duality matrix $D$ and we have
\begin{align}\label{useful-InInterpretation}
	\Psi^{'}(\bm{\tau})=&\sum_{\sigma\in \Omega}\langle \bm{\sigma}|D|\bm{\xi}(\bm{\tau})\rangle\langle \bm{\sigma}|\Psi\rangle\dt
\end{align}
We observe that for any $x\in \{1,\ldots,L\}$
\begin{equation}
	\prod_{a=1}^{N-1}\mathbbm{1}_{\{\delta_{\sigma_{x},a}\geq\xi_{a}^{x}(\bm{\tau})\}}=\left(\mathbbm{1}_{\{\sigma_{x}=\tau_{x}\}}\right)^{(1-\delta_{\tau_{x},N})}%=\begin{cases}
	%	\mathbbm{1}_{\{\sigma_{x}=\tau_{x}\}}\qquad &\text{if}\quad \tau_{x}\neq N\\
	%	1\qquad  &\text{if}\quad \tau_{x}= N
%	\end{cases}\dt
\end{equation}
Indeed, we have that 
\begin{equation}
	\prod_{a=1}^{N-1}\mathbbm{1}_{\{\delta_{\sigma_{x},a}\geq\xi_{a}^{x}(\bm{\tau})\}}=\begin{cases}
			\mathbbm{1}_{\{\sigma_{x}=\tau_{x}\}}\qquad &\text{if}\quad \tau_{x}\neq N\\
			1\qquad  &\text{if}\quad \tau_{x}= N
			\end{cases}\dt
\end{equation}
 Therefore using the duality function \eqref{ElementsDualityMatrixChain_1} and equation~\ref{useful-InInterpretation}, we obtain 
 	\begin{equation}\label{interpretation-PsiP}
 	\Psi^{'}(\bm{\tau})=\mathbb{E}\left[\prod_{x=1}^{L}\left(\mathbbm{1}_{\{Y_{x}=\tau_{x}\}} \right)^{(1-\delta_{\tau_{x},N})}\right]\dt
 \end{equation} 
 This last equation completes and generalizes \eqref{corr-psi1} by showing that $\Psi^{'}(N,\ldots,N) =1$.

@@

@@

The computations made for the multispecies stirring process can be specialized to the single-species case $N=2$. In this situation we retrieve the SSEP, that was already solved with the quantum inverse scattering method in \cite{frassek2020eigenstates}. We report the Hamiltonians and the similarity transformations of SEEP: 

@@

@@

$1\leq \ell_{1}<\ell_{2}<\ldots<\ell_{q}\leq m$. To each of these label we associate the coordinates $x_{\ell_{1}},\ldots,x_{\ell_{q}}\in\{x_{1},\dots,x_{m}\}$ that are such $c_{x_{\ell_{k}}}=1$ for all $k\in \{1,\ldots,q\}$ and $c_{x}=0$ for $x\notin \{x_{\ell_{1}},\ldots,x_{\ell_{q}}\}$. Therefore, for any fixed $c_{x_{1}},\ldots,c_{x_{m}}$, we find $q$ and $\ell_{1},\ldots,\ell_{q}$ from which we fix $x_{\ell_{1}},\ldots,x_{\ell_{q}}$. Then, we obtain  

>>>>

Observing now that $c_{x}(1-\delta_{\tau_{x},N})$ can be replaced by  $c_{x}-\delta_{\tau_{x},N}$, we finally have
\begin{equation}\label{elementsABS}
	\begin{split}
		\Psi^{'}(\bm{\tau})=\frac{1}{(L+1)!}\sum_{c_{1}=\delta_{\tau_{1},N}}^{1}\ldots\sum_{c_{L}=\delta_{\tau_{L},N}}^{1}\prod_{x=1}^{L}(\alpha_{\tau_{x}}-\beta_{\tau_{x}})^{1-c_{x}}\left(1+\sum_{j=x}^{L}c_{j}\right)\beta_{\tau_{x}}^{c_{x}-\delta_{\tau_{x},N}}\dt
	\end{split}
\end{equation} 
\begin{flushright}
    $\square$
\end{flushright}
\begin{remark} By changing the summation indices from $c_x\in \{\delta_{\tau_{x},N},1\}$ to $c_{x}^{'}\in \{0,1-\delta_{\tau_{x},N}\}$ and by  the fact that occupation of each site is bounded by $1$ we rewrite \eqref{ABS} as 
\begin{equation}\label{ABS_1}
	\Psi^{'}(\bm{\tau})=\sum_{c^{'}_{1}=0}^{1-\delta_{\tau_{1},N}}\ldots\sum_{c^{'}_{L}=0}^{1-\delta_{\tau_{L},N}}\frac{(L+1-\sum_{x=1}^{L}c^{'}_{x})!}{(L+1)!}\prod_{x=1}^{L}\left((\alpha_{\tau_{x}}-\beta_{\tau_{x}})\left(2+L-x-\sum_{j=x}^{L}c^{'}_{j}\right)\right)^{c^{'}_{x}}\beta_{\tau_{x}}^{(1-c^{'}_{x}-\delta_{\tau_{x},N})}
\end{equation}
where we used the fact that 
\begin{equation}
	\left(1+\sum_{j=x}^{L}c_{j}\right) =\left(2+L-x-\sum_{j=x}^{L}c^{'}_{j}\right)\dt
\end{equation}
Equation \eqref{ABS_1} will be useful in the following. 
\end{remark}
\subsubsection{Closed formula for correlations}\label{correlation-section}
We consider $m\in \{1,\ldots,L\}$ sites with coordinates $x_{1},\ldots,x_{m}\in \{1,\ldots,L\}$ such that $x_{k}<x_{k+1}$ for all $k\in \{1,\ldots,m-1\}$. We fix $a_{k}\in\{1,\ldots,N-1\}$ and we would like to compute
\begin{equation}
	\mathbb{E}\left[\prod_{k=1}^{m}\mathbbm{1}_{\{Y_{x_{k}}=a_{k}\}}\right]=\sum_{\bm{\tau}\in \Omega^{'}}\left(\prod_{k=1}^{m}\mathbbm{1}_{\{\tau_{x_{k}}=a_{k}\}}\right)\langle\bm{\tau}|\Psi\rangle\dt%=\sum_{\bm{\tau}\in \Omega^{'}}\left(\prod_{i=1}^{m}\mathbbm{1}_{\{\tau_{y_{i}}=s_{i}\}}\right)\Psi(\bm{\tau})
\end{equation}
We fix the dual configuration $\bm{\xi}=\sum_{k=1}^{m}\delta_{a_{k}}^{x_{k}}$, corresponding to a vector $|\bm{\xi}\rangle$, and we have 
\begin{equation}
	\prod_{k=1}^{m}\mathbbm{1}_{\{\tau_{x_{k}}=a_{k}\}}=D(\bm{\tau},\bm{\xi})\dt
\end{equation}
Thus it follows that 
\begin{align}
	\mathbb{E}\left[\prod_{k=1}^{m}\mathbbm{1}_{\{Y_{x_{k}}=a_{k}\}}\right]=&\sum_{\bm{\tau}\in \Omega^{'}}D(\bm{\tau},\bm{\xi})\langle\bm{\tau}|\Psi\rangle
	=\sum_{\bm{\tau}\in \Omega^{'}}\langle \bm{\tau}|D|\bm{\xi}\rangle\langle\bm{\tau}|\Psi\rangle\dt
\end{align}
Denoting by $|\hat{\bm{\xi}}\rangle=|\xi^{1},\ldots,\xi^{L}\rangle$, i.e. the vector constructed from $|\bm{\xi}\rangle$ by removing the components at sites $\{0\}$ and $\{L+1\}$ and recalling that  
\begin{equation}
	S_{1}^{T}=\prod_{x=1}^{L}d_{x}
\end{equation} 
\todo{double check and ref to $d_x$}
we have
\begin{equation}
	\langle \bm{\tau}|S_{1}^{T}|\hat{\bm{\xi}}\rangle=\langle \bm{\tau}|D|\bm{\xi}\rangle \dt
\end{equation}
Therefore, it follows that, using the resolution of the identity
\begin{align}
\label{corr-psi1}
	\mathbb{E}\left[\prod_{k=1}^{m}\mathbbm{1}_{\{Y_{x_{k}}=a_{k}\}}\right]=& \sum_{\bm{\tau}\in \Omega^{'}}\langle \bm{\tau}|D|\bm{\xi}\rangle\langle\bm{\tau}|\Psi\rangle
	=\sum_{\bm{\tau}\in \Omega^{'}}\langle \bm{\tau}|S_{1}^{T}|\hat{\bm{\xi}}\rangle\langle\bm{\tau}|\Psi\rangle\nonumber\\
	=&\sum_{\bm{\tau}\in \Omega^{'}}\langle \hat{\bm{\xi}}|S_{1}|\bm{\tau}\rangle\langle\bm{\tau}|\Psi\rangle
	=\langle \hat{\bm{\xi}}|S_{1}|\Psi\rangle
	=\langle \hat{\bm{\xi}}|\Psi^{'}\rangle\dt
\end{align}

@@

{\color{blue}
\begin{remark} 
Without using duality but only  the definition $|\Psi^{'}\rangle=S_1|\Psi\rangle$ and that
\begin{equation}
 \langle \tau |S_1=\begin{cases}
                    \sum_{\sigma=1}^N\langle \sigma|  \quad\text{if}\quad \tau=N\\
                    \langle \tau| \quad\text{else} 
                   \end{cases}\,,
\end{equation} 
  we immediately verify 
\begin{equation}\label{eq:EE}
\begin{split}
\mathbb{E}\left[\prod_{k=1}^{m}\mathbbm{1}_{\{Y_{x_{k}}=a_{k}\}}\right]=\sum_{\bm{\tau}\in\Omega^{'}}\prod_{k=1}^{m}\delta_{{{\tau}_{x_k}},a_{k}}\langle \bm{\tau}|\Psi\rangle=
\langle N,\ldots,N,a_1,N,\ldots,N,a_m,N,\ldots,N|\Psi^{'}\rangle\dt
 \end{split}
\end{equation} 
Here the entries $a_i\neq N$ in \eqref{eq:EE} are at position $x_i$. 
\end{remark}

}
\todo{ok? adopt notation?}

\subsubsection{Interpretation of $\Psi^{'}(\bm{\tau})$ as marginals of the steady state distribution}\label{subsection-interpretationPSIp}
\todo{discuss section}
%We argue similarly as in the Section~\ref{correlation-section}. 
We still miss the interpretation of $\Psi^{'}(\bm{\tau})$ when  $\bm{\tau}= (N,\ldots,N)$. We show here that for this particular configuration
$\Psi^{'}(N,\ldots,N) =1$. Indeed, for any $\bm{\tau}\in \Omega^{'}$ we have that 
\begin{align}
	\Psi^{'}(\bm{\tau})=&\langle\bm{\tau}|\Psi^{'}\rangle=\sum_{\bm{\sigma}\in \Omega}\langle \bm{\tau}|S_{1}|\bm{\sigma}\rangle\langle \bm{\sigma}|\Psi\rangle=\sum_{\sigma\in \Omega}\langle \bm{\sigma}|S_{1}^{T}|\bm{\tau}\rangle\langle \bm{\sigma}|\Psi\rangle\dt
\end{align}
\todo{what was Omega again?}
We introduce the dual configuration $\bm{\xi}(\bm{\tau})=(N,\tau_{1},\ldots,\tau_{L},N)$ where two extra-sites occupied by the species $N$ have been attached to the vector $\bm{\tau}$, transforming it from a vector of dimension $L$ to one of dimension $L+2$. More explicitly, this dual configuration is given by
\begin{equation}
	\begin{cases}
	\xi_{a}^{x}(\bm{\tau})= \delta_{\tau_{x},a}\qquad \forall a\in \{1,\ldots N\}\quad \forall x\in \{1,\ldots,L\}\\
	\xi_{N}^{0}(\bm{\tau})=\xi_{N}^{L+1}(\bm{\tau})=1
\end{cases}\dt
\end{equation} Now we consider the elements $\langle\bm{\sigma}|D|\xi(\bm{\tau})\rangle$ of the duality matrix $D$ and we have
\begin{align}\label{useful-InInterpretation}
	\Psi^{'}(\bm{\tau})=&\sum_{\sigma\in \Omega}\langle \bm{\sigma}|D|\bm{\xi}(\bm{\tau})\rangle\langle \bm{\sigma}|\Psi\rangle\dt
\end{align}
We observe that for any $x\in \{1,\ldots,L\}$
\begin{equation}
	\prod_{a=1}^{N-1}\mathbbm{1}_{\{\delta_{\sigma_{x},a}\geq\xi_{a}^{x}(\bm{\tau})\}}=\left(\mathbbm{1}_{\{\sigma_{x}=\tau_{x}\}}\right)^{(1-\delta_{\tau_{x},N})}%=\begin{cases}
	%	\mathbbm{1}_{\{\sigma_{x}=\tau_{x}\}}\qquad &\text{if}\quad \tau_{x}\neq N\\
	%	1\qquad  &\text{if}\quad \tau_{x}= N
%	\end{cases}\dt
\end{equation}
Indeed, we have that 
\begin{equation}
	\prod_{a=1}^{N-1}\mathbbm{1}_{\{\delta_{\sigma_{x},a}\geq\xi_{a}^{x}(\bm{\tau})\}}=\begin{cases}
			\mathbbm{1}_{\{\sigma_{x}=\tau_{x}\}}\qquad &\text{if}\quad \tau_{x}\neq N\\
			1\qquad  &\text{if}\quad \tau_{x}= N
			\end{cases}\dt
\end{equation}
 Therefore using the duality function \eqref{ElementsDualityMatrixChain_1} and equation~\ref{useful-InInterpretation}, we obtain 
 	\begin{equation}\label{interpretation-PsiP}
 	\Psi^{'}(\bm{\tau})=\mathbb{E}\left[\prod_{x=1}^{L}\left(\mathbbm{1}_{\{Y_{x}=\tau_{x}\}} \right)^{(1-\delta_{\tau_{x},N})}\right]\dt
 \end{equation} 
 This last equation completes and generalizes \eqref{corr-psi1} by showing that $\Psi^{'}(N,\ldots,N) =1$.

@@

		\todo{can we remove q index in sum?}

@@

The computations made for the multispecies stirring process can be specialized to the single-species case $N=2$. In this situation we retrieve the SSEP, that was studied using the same approach as presented here in \cite{frassek2020eigenstates}. We report the Hamiltonians and the similarity transformations of SEEP: 

@@

\todo{double check}

@@

$1\leq \ell_{1}<\ell_{2}<\ldots<\ell_{q}\leq m$. 
To each of these label we associate the coordinates $x_{\ell_{1}},\ldots,x_{\ell_{q}}\in\{x_{1},\dots,x_{m}\}$ that are such $c_{x_{\ell_{k}}}=1$ for all $k\in \{1,\ldots,q\}$ and $c_{x}=0$ for $x\notin \{x_{\ell_{1}},\ldots,x_{\ell_{q}}\}$. Therefore, for any fixed $c_{x_{1}},\ldots,c_{x_{m}}$, we find $q$ and $\ell_{1},\ldots,\ell_{q}$ from which we fix $x_{\ell_{1}},\ldots,x_{\ell_{q}}$. Then, we obtain  
